Viscous Transport in Eroding Porous Media

### 2019-01-25 docs/transport/erosion_transport.tex
started to lay out transport methods paper

\title{A boundary-integral framework to simulate \\ viscous erosion of a porous medium}

@@

\author[Bryan]{Bryan D.~Quaife}
\author[Nick]{M.~Nicholas J.~Moore}
\address[Nick]{Department of Mathematics and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\address[Bryan]{Department of Scientific Computing and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}

\begin{abstract} 
We develop numerical methods to simulate the fluid-mechanical erosion of many bodies in two-dimensional Stokes flow. The broad aim is to simulate the erosion of a porous medium (e.g.~groundwater flow) with grain-scale resolution. Our fluid solver is based on a second-kind boundary integral formulation of the Stokes equations that is discretized with a spectrally-accurate Nystr\"om method and solved with fast-multipole-accelerated GMRES.  The fluid solver provides the surface shear stress which is used to advance solid boundaries. We regularize interface evolution via curvature penalization using the {\thL} formulation, which affords numerically stable treatment of stiff terms and therefore permits large time steps.  The overall accuracy of our method is spectral in space and second-order in time. The method is computationally efficient, with the fluid solver requiring $\mathcal{O}(N)$ operations per GMRES iteration, a mesh-independent number of GMRES iterations, and a one-time $\mathcal{O}(N^2)$ computation to compute the shear stress. We benchmark single-body results against analytical predictions for the limiting morphology and vanishing rate. Multibody simulations reveal the spontaneous formation of channels between bodies of close initial proximity. The channelization is associated with a dramatic reduction in the resistance of the porous medium, much more than would be expected from the reduction in grain size alone.
\end{abstract}

\begin{keyword}
  Porous medium \sep Stokes flow \sep Erosion \sep Boundary integral
  equations \sep Shear stress \sep {\thL} formulation
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from massive land formations~\cite{han1969, perron2009formation, abrams2009growth, petroff2011geometry, jerolmack2012internal, Rothman2012, coh-dev-sey-yi-szy-rot2015, perkins2015amplification}, to centimeter-scale features and patterns~\cite{daerr2003erosion, berhanu2012shape, nienhuis2014wavelength, domokos2014river}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~soil, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in Stokes flow---the most relevant regime for groundwater \cite{bear2013dynamics, cao2010coupled} and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann, lopez2018cfd} and the related processes of dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, dallaston2015channelization, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the fluid alters the morphology of immersed structures, which in turn modifies the surrounding flow field, and so on. Such examples of `sculpting with flow'~\cite{ristroph2018sculpting} have been examined in the context of a single body~\cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shape evolution is intertwined~\cite{hewett2017evolution}. From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multiple bodies must be considered. Second, the dense packing present in typical porous-media applications requires close-range interactions to be resolved. Finally, long-range effects inherent to the Stokes limit couples the shape evolution of {\em all} bodies to a greater degree than any other regime.

Ultimately, the methods developed here will lead to high-fidelity simulations of multibody Stokes erosion as illustrated in Fig.~\ref{fig:50bodies} for the case of 50 bodies (of randomly assigned initial sizes and positions). Here, the erosion of solid material arises from the mechanical stress induced by the surrounding flow. In particular, we take the local erosion rate to be proportional to the magnitude of the shear stress---a law that has been verified by laboratory experiments~\cite{ris-moo-chi-she-zha2012}. As seen in Fig.~\ref{fig:50bodies}, erosion not only diminishes the size of the bodies, but also alters their shapes considerably. In particular, several bodies develop corner-like features and nearly flat faces, with straight channels intervening between them. In the figure, color represents the vorticity of the surrounding flow, which reduces to shear on solid boundaries and thus provides a convenient way to visualize local erosion rates. Several bodies vanish in finite time and the simulation continues without interruption.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.80 \textwidth]{./figs/50bod.pdf}
\caption{\label{fig:50bodies} Simulation of 50 bodies eroding in Stokes flow under the action of fluid-mechanical shear. The flow is horizontal (left to right) and the 6 snapshots are evenly spaced in time. Color represents vorticity, which provides a convenient way to visualize local shear rates. Erosion not only diminishes the size of the bodies but also alters their shapes considerably. A few well-defined channels develop in the space between bodies. Several of the bodies vanish in finite time and the simulation continues without interruption.
}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
To perform the above simulation, two fast and accurate methods are
employed---one to solve the fluid equations and the other to evolve the
eroding bodies.  The fluid equations are reformulated in terms of a boundary integral equation (BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we employ a double-layer potential so that the resulting BIE is second-kind.  This second-kind BIE is discretized with a spectrally accurate trapezoid rule~\cite{tre-wei2014} which results in a well-conditioned dense linear system for an unknown density function. The linear system is solved using fast-multipole-accelerated GMRES~\cite{saa-sch1986}.  Once the density function is determined, we compute the shear stress with a spectrally accurate odd-even quadrature rule~\cite{sid-isr1988}.

At the next stage, we must use the shear stress to advance solid boundaries. Tracking fluid-structure interfaces with long time horizons presents numerical challenges in that the underlying mesh can become distorted and tangled. Furthermore, the process of erosion naturally leads to regions of high curvature and even corners.  We address these tracking issues in two ways: first, we introduce a tangential velocity that keeps the grid equispaced with respect to arclength but that does not affect dynamics; second, we suppress high-frequency oscillations in the interface evolution via curvature penalization. The interface is parametrized in the so-called {\thL} coordinate system so that the curvature penalization appears as a diffusive term \cite{hou-low-she1994}. This term can be easily be handled in a numerically stable fashion with an exponential integrator.  Once the simulation is complete, we perform a post-processing step in which we compute bulk properties, such as vorticity and pressure, using an upsampled trapezoid rule.

We supplement our primarily computational study with a few analytical results. In particular, our simulation shows that a single eroding body tends to a slender, for-aft symmetric form characterized by nearly uniform shear stress along its surface. This body exhibits sharp corners at its front and rear. The principle of long-time uniform stress combined with some local analysis leads to an analytical prediction for the opening angles formed at the front and rear. Additionally, slender-body theory gives rise to scaling laws for the reduction rate of the body's area and drag. These analytical results, in addition to their own intrinsic value, allow us to conclusively validate the numerical simulation.

\paragraph{Limitations}
We make two fundamental design choices in the simulations described here: first, we restrict attention to two spatial dimensions, and second the bodies are considered immobile (i.e.~fixed to their initial locations). By restricting to two dimensions, we can simulate many bodies (up to $50$ in this manuscript) with high resolution, enabling accurate characterization of porous-medium erosion. While BIEs are available in three dimensions, the extra computational cost limits the number of bodies that could be practically simulated and makes porous-medium simulations less feasible. Meanwhile, the second assumption of immobile bodies affords a simple framework to explore the effects of erosion in isolation. While other mechanisms, such as sedimentation or transport, undoubtedly play a role in real applications, these effects would introduce additional complexity, both conceptually and numerically. We therefore defer their study to future work.

\paragraph{Related Work}
In a closely related study, Mitchell \& Spagnolie used a traction integral equation formulation to simulate the viscous erosion of a single body in three dimensions~\cite{mit-spa2016}. The flexibility of this formulation enabled simulations of a {\em mobile} body eroding in an {\em arbitrary} background flow (e.g.~uniform flow, shear flow, etc.).  An advantage of the traction formulation is that it directly provides the shear stress needed to evolve surfaces.  In the present work, we aim to simulate the erosion of a porous medium and, further, choose to fix the bodies to their initial locations. Thus, rather than a {\em single, mobile} body as considered in~\cite{mit-spa2016}, we aim to simulate the erosion of {\em many, immobile} bodies. Although the traction formulation has several attractive features, we instead employ a standard double-layer formulation since it allows us to easily calculate bulk quantities (e.g.~vorticity, pressure, and velocity) that will assist our understanding of the erosion process.

For the porous network simulations considered here, it is essential to
resolve nonlocal effects between the different bodies.  BIEs naturally
resolve these interactions, and they have been used extensively to
simulate viscous fluids in complex geometries, including porous media
flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2018}.  Common challenges
of a BIE formulation include developing efficient preconditioners for
the discretized system and numerical difficulties associated with nearly
touching bodies. In this work, we consider geometries requiring
$\mathcal{O}(100)$ GMRES iterations, so preconditioners are currently
unnecessary. In future work, we will address mobile bodies for which
near contacts may necessitate preconditioners~\cite{qua-bir2015a, qua-cou-dar2018, cou-pou-dar2017}. Similarly, near-singular schemes will not be necessary due to the fact that the spacing between receding bodies of fixed position always grows in time. Again, the future consideration of mobile bodies may necessitate near-singular schemes~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015, hel-oja2008a, bea-lai2001}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{s:formulation}, we describe the governing equations for both the fluid and the erosion model in a \thL~framework.  In Section~\ref{s:method}, the numerical methods in both space and time are described and Section~\ref{s:qoi} describes how to compute the vorticity, pressure, and drag.  Section~\ref{s:SingleResults} gives several numerical examples for a single body, and Section~\ref{s:MultiResults} presents numerical examples for multiply-connected domains.  Finally, concluding remarks are made in Section~\ref{s:conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We consider flows that are confined by a solid wall $\Gamma$ that encloses $M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  Therefore, the boundary of the fluid domain is $\bd\Omega = \Gamma \cup \gamma$. On each body, $\nn$ and $\ss$ denote the unit normal and unit tangent vectors respectively.  Throughout the paper, we adopt the convention that $\ss$ points in the counterclockwise direction and $\nn$ points out of the fluid domain (i.e.~into the bodies). Neglecting inertial forces, the dynamics of the fluid is fully determined by the position of the bodies $\xx_\ell(s,t) \in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given $\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model} .
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface. Parameters include the fluid viscosity $\mu$ and the material-dependent erosion constant $\CE$, as well as a characteristic flow speed $U^*$ and domain scale $L^*$. Non-dimensionalization, however, allows us to remove these parameters from consideration. We non-dimensionalize the Stokes equations in the standard way, equivalent to setting $\mu = 1$ above. For interface evolution, we scale time on the so-called erosion timescale $T^* = {L^*}^2/(\CE \mu U^*)$, which can be achieved by simply setting $\CE = 1$ above. In Section~\ref{sec:thetaL}, we will make slight modifications to the erosion model for better numerical stability.

The outer wall is formed by rounding off the corners of $[-3,3] \times [-1,1]$, and we impose the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx) = \umax (1-y^2,0)\, , \quad \xx \in \Gamma\, ,
\end{align}
where $U$ (dimensionless) sets the maximum of the imposed Poiseuille flow. Often, we will simply set $U=1$, but in some instances we change $U$ dynamically to enforce that the pressure drop across the computational cell is constant as bodies erode within it. To minimize the boundary effects of the inflow and outflow, we only place bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing equations.  A no-slip boundary condition is imposed on each body $\gamma_\ell$ whose unit normal points outward relative to the geometry.  On the outer geometry $\Gamma$, a Hagen-Poiseuille flow is imposed.  The bodies erode at a rate that is equal to the magnitude of the shear stress applied by the fluid.  The bodies are constrained to the middle third of the channel that is located between the dashed lines.}
\end{figure}

We describe how the incompressible Stokes equations are solved in $\Omega$ using a boundary integral equation in Section~\ref{sec:bies}.  In Section~\ref{sec:shearStressLP}, we discuss how the shear stress is computed.  Finally, in Section~\ref{sec:thetaL}, we describe how corners in the bodies are avoided and perform a change of coordinates to treat a stiff term.

@@

In this section, we describe an equivalent BIE formulation of the incompressible Stokes equations with Dirichlet boundary conditions
\begin{equation}
  \begin{split}
  \Delta \uu &= \grad p\, , \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
  \end{split}
  \label{eqn:stokes}
\end{equation}
The boundary condition $\ff$ is the prescribed velocity $\UU$ on
$\Gamma$, and it is zero on the eroding bodies $\gamma$.  In two dimensions, a BIE for a stream function formulation is possible~\cite{gre-kro-may1996}, but this does not extend to three dimensions.  Therefore, we adopt a primitive variable formulation where the velocity is written in terms of the standard double-layer potential~\cite{lad1963,poz1992}
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy\, , \quad \xx \in \Omega \, ,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\eeta$ is an unknown density function.  
%As the viscosity $\mu$ also sets the time scale, we take $\mu=1$.

Since the double-layer potential is not able to represent force- and torque-free velocity fields,  we complete the double-layer potential with Stokeslets and rotlets~\cite{pow-mir1987, pow1993}
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell\, , 
  \quad \ell=1,\ldots,M \, ,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the $\ell^{th}$ grain, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then, the solution of~\eqref{eqn:stokes} is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx)\, , \quad \xx \in \Omega \, ,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx)\, ,
      \quad \xx \in \bd\Omega \, , 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \,
  ds_\yy\, ,
  \quad \ell=1,\ldots,M \, ,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy\, , \quad \ell=1,\ldots,M \, .
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy, \quad \xx \in
    \Gamma\, ,
\end{align}
removes the rank one null space resulting from the flux-free requirement of $\ff$.

This BIE formulation requires that $\bd\Omega$ is smooth.  Unfortunately, smoothness is not guaranteed by erosion since corners develop at locations where the shear stress vanishes.  One option is to modify our BIE to handle corners~\cite{rac-ser2017, ser-rok2016, hel2011, gil-hao-mar2014, bre2012}, but these methods require the corner's location, a non-uniform spaced mesh, or are ill-conditioned.  Instead, we slightly modify the erosion model in Section~\ref{sec:thetaL} so that we can control the smoothness of the bodies.  In this manner, the BIE as described above can be applied.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the shear stress}
\label{sec:shearStressLP}
Once the density function, Stokeslets, and rotlets are computed, any desired quantity, including the shear stress, can be computed.  The shear stress is
\begin{align}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss \, .
\end{align}
The reader may notice an unusual sign convention above, but this is simply a consequence of our conventions for $\nn$ and $\ss$ (see Fig.~\ref{fig:schematic}).  The velocity field of the completed double-layer potential~\eqref{eqn:completed_DLP} includes velocities from three terms: the double-layer potential, the Stokeslets, and the rotlets.  We compute the deformation tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of these terms individually.

The deformation tensor of the double-layer potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$ is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \, ds_\yy \, .
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0)\, , \quad \xx_0 \in \bd\Omega \, ,
\end{align} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] \, .
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right) \, .
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss \, .
  \label{eqn:shearStressLP}
\end{align}

@@

To evolve the boundaries of eroding bodies, we use the {\thL} framework which offers certain advantages in the ability to numerically stabilize moving-interface problems~\cite{hou-low-she1994}. Here, $\theta$ is the local tangent angle and $L$ is the total perimeter of a given boundary. For a boundary parametrized by arclength $(x(s),y(s))$, the local tangent angle is defined through the relation $\left( \pdi{x}{s}, \pdi{y}{s} \right) = \left(\cos \theta, \sin \theta \right)$.

For numerical stability, we modify the erosion law~\eqref{eqn:erosionModel} to include a term that penalizes curvature $\kappa$. The idea is to prevent regions of extremely high curvature from developing, as these regions would compromise the accuracy and stability of our fluid solver. We thus replace~\eqref{eqn:erosionModel} with
\begin{align}
  \label{eqn:Vn}
  \Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1
  \right) \, .
\end{align}
The parameter $\eps \ll 1$ controls the strength of the curvature-penalization. This term causes regions of high curvature to recede more rapidly and thus suppresses any high-frequency oscillations that may exist. Above, $\mean{\cdot}$ indicates a boundary mean. Since a closed body has mean curvature $\mean{\kappa} = 2\pi / L$, the penalty term has mean zero and thus preserves area. The only source of material loss is therefore the shear-stress itself. In addition, we have scaled the penalty term on the mean absolute shear, $\mean{\atau}$, so that the ratio of smoothing-to-erosion always remains fixed, even as bodies vanish.

In the following, we parameterize each boundary by normalized arclength, $\alpha = 2 \pi s / L$. In terms of $\alpha$, the curvature is given by
\begin{align}
\kappa = \frac{2 \pi}{L} \pderiv{\theta}{\alpha} \, ,
\end{align}
and substitution into~\eqref{eqn:Vn} gives
\begin{align}
\label{eqn:Vn2}
  \Vn = \atau +  \eps \mean{\atau}   \left(\pderiv{\theta}{\alpha} - 1 \right) \, .
\end{align}

The boundary of each body will be represented by a grid of collocation points. If we were to simply evolve the interfaces under~\eqref{eqn:Vn2}, this grid would become distorted, which, if unchecked, could lead to severe tangling and numerical instability. We therefore introduce an artificial tangential velocity $\Vs$ to keep the grid points equally spaced in arclength~\cite{hou-low-she1994}. The appropriate tangential velocity satisfies 
\begin{align}
  \pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn} \, .
\end{align} 
The evolution of the interface is then given by the combination of the normal and tangential velocities
\begin{align}
  \dot{\xx}(t) = \Vn \nn + \Vs \ss \, .
\end{align}
Since $\Vs$ is tangential, it does not modify the physics of erosion, only the parameterization of the boundaries.

When recast in the variables $\theta$ and $L$, the interface evolution becomes
\begin{align}
  \tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
  \pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} + \thalpha \Vs \right) \, .
\end{align}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
  \pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
  \eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
  \thalpha \Vs \right)\,.
\end{align}
These are the main evolution equations for $\theta(\alpha,t)$ and $L(t)$.  In addition, it is necessary to track a reference point for each body, so that the {\thL} variables can be converted back to Cartesian coordinates. We track the coordinates of the surface mean $\mean{\xx}$, which moves according to
\begin{align}
\label{eq:xsurf}
\frac{d}{dt} \mean{\xx} = \mean{ V_s \ss + V_n \nn } \, .
\end{align}
We note that tracking the center of mass is also a valid choice, but the surface mean gives a somewhat simpler formula. Equations~\eqref{eqn:Levo1}--\eqref{eq:xsurf} comprise the main system of equations to be solved numerically in Section~\ref{sec:timeStepping}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}
To perform numerical simulations, we exploit the difference between the erosion and fluid time scales.  Since erosion occurs much slower, we freeze the geometry at each time step, solve the incompressible Stokes equations, compute the shear stress, and then update the geometry according to the erosion model.  Therefore, there are three main numerical methods that need to be developed.  In Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE} is solved.  Once the density function is found, Section~\ref{sec:shearStress} describes how the shear stress~\eqref{eqn:shearStressLP} is computed.  Then, Section~\ref{sec:timeStepping} describes a time stepping method for advancing the interface according to equations~\eqref{eqn:Levo1}--\eqref{eq:xsurf}.

To create stable simulations with long time horizons, we use numerical methods that achieve spectral accuracy in space, second-order accuracy in time, and maintain an equispaced discretization of the bodies.  The boundaries are represented in Fourier space, giving us a set of discretization points $\xx_i$ of each body, and all derivatives and integrals (the arclength, curvature, and mean values) are computed with spectral accuracy using the Fourier representation.

@@

With the shear stress computed, we are now ready to numerically evolve the boundary of each body using the {\thL} framework. The main equations to be solved are~\eqref{eqn:Levo1}--\eqref{eq:xsurf}.  In Section~\ref{sec:thetaL} we introduced a curvature-penalizing term into the erosion law~\eqref{eqn:Vn}. This term suppresses high-frequency oscillations, but also introduces {\em stiffness} into evolution equation~\eqref{eqn:thetaevo1}. We will therefore use a combined exponential-integrator/Runge-Kutta method to solve system~\eqref{eqn:Levo1}--\eqref{eq:xsurf}. The exponential integrator will handle the stiff term in a numerically stable fashion, while the Runge-Kutta method will treat nonlinear terms explicitly. Our overall time-stepping method will be second-order accurate.

First, we must address another potential source of numerical instability. The erosion law~\eqref{eqn:Vn} depends on the {\em absolute value} of shear stress, $\atau$, which exhibits a corner anywhere $\tau$ changes sign, i.e.~at a stagnation point. The corner in $\atau$ introduces high frequencies into~\eqref{eqn:thetaevo1} which could result in either numerical instability or an extremely stringent step-size restriction. To avoid these difficulties, we smooth $\atau$ via a narrow Gaussian filter of width $\sigma \ll 1$
\begin{align}
\atausig = \int_{\gamma} \frac{1}{\sqrt{2\pi} \sigma}
 \exp \left(- \frac{\alpha'^2}{2 \sigma^2}\right) \abs{ \tau(\alpha - \alpha') } \, d\alpha' \, ,
\end{align}
with corresponding normal and tangential interface velocities
\begin{align}
& \Vnsig = \atausig +  \eps \mean{\atausig}
\left(\pderiv{\theta}{\alpha} - 1 \right), \\
& \pderiv{\Vssig}{\alpha} = \thalpha \Vnsig - \mean{\thalpha \Vnsig} \, .
\end{align}
This smoothing further suppresses the feedback of high frequencies between $\atau$ and the body shape.

>>>>

\title{Viscous Transport in Eroding Porous Media}

@@

\author[SH]{Shang-Huan Chiu}
\author[Nick]{M.~Nicholas J.~Moore}
\author[Bryan]{Bryan D.~Quaife}

\address[SH]{Department of Scientific Computing, Florida State
University, Tallahassee, FL, 32306.}
\address[Nick]{Department of Mathematics and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}
\address[Bryan]{Department of Scientific Computing and Geophysical Fluid Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}

\begin{abstract} 
  An abstract
\end{abstract}

\begin{keyword}
  Keyword 1 \sep Keyword 2 \sep Keyword 3 
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
A continuation of the methods paper~\cite{qua-moo2018}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
\begin{itemize}
  \item Use novel near-singular Barycentric integration scheme for
    Stokes equation outlined in~\cite{bar-wu-vee2015}.
  \item Need to compute one additional derivative since we need the
    derivative of the velocity to find the vorticity and stress tensor
  \item Tracer trajectories at different time steps
  \item Compute the anomalous diffusion rate using similar techniques as
    the paper with Pietro~\cite{dea-qua-bir-jua2018}
  \item Need a novel way to do time stepping through the complex
    geometry
\end{itemize}

\paragraph{Limitations}
\begin{itemize}
  \item Bodies are still pinned
  \item Bodies are still two dimensional
\end{itemize}

\paragraph{Related Work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
Similar to the JCP methods paper.  However, need to put everything into
complex variables formulation so that we can use Barycentric ideas.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the shear stress}
\label{sec:shearStressLP}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation Methods}
\label{sec:FMM}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Post processing quantities of interest}
\label{s:qoi}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vorticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pressure}
\label{sec:pressure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Drag}
\label{sec:drag}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Near-singular integration}
\label{sec:NSI}

%%-----------------------------------------------------------------%%
% RESULTS SECTION
%%-----------------------------------------------------------------%%
\section{Results: Single-body erosion}
\label{s:SingleResults}
